### 2023-06-16 tesis_Ashish/Portada/Portada.tex
Your commit message

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%
%
%  ESPACIO COMENTADO PARA EVITAR DESAJUSTES POR GENERACIÓN DE NUEVO PÁRRAFO (NO DEJAR ESPACIOS ENTRE MINIPAGES)
%
%
%
%
%
%
%%%%%%%%%%%%%%%% 5TO MINIPAGE PARA EL CUERPO DE LA PORTADA (NOMBRES, TÍTULO, ETC.) %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%
%\fbox{
\begin{minipage}[l][0.78\textheight][t]{0.1\textwidth}
    \begin{center}
    \hskip0pt
    \vrule width2.5pt height17cm    %% height definde la altura de la linea gruesa
        \hskip1mm
        \vrule width1pt height17cm  %% height define la altura de la linea delgada
        \end{center}
\end{minipage}
%}
%\fbox{
\begin{minipage}[c][0.78\textheight][t]{0.8\textwidth}
      \begin{center}
      \vspace{2cm}
        {\Large \scshape {Measurements of the Large Hadron Collider luminosity using the CMS silicon pixel detector}}

@@

        \makebox[5cm][c]{\LARGE THESIS}  \\[20pt]
         in partial fulfillment of the
requirements for the degree of:\\[5pt]
        {\Large \textbf{{Maestr\'ia en Ciencias (F\'isica)}}}\\[40pt]
        By:\\[12pt]
        \textbf{  {\large Luis Enrique CUEVAS PICOS}}

@@

        { Hermosillo, Sonora} \hskip2.5cm {May, 2023}
      \end{center}
\end{minipage}
%}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%
%
%  ESPACIO COMENTADO PARA EVITAR DESAJUSTES POR GENERACIÓN DE NUEVO PÁRRAFO (NO DEJAR ESPACIOS ENTRE MINIPAGES)
%
%
%
%
%
%
%%%%%%%%%%%%%%%% 5TO MINIPAGE PARA EL CUERPO DE LA PORTADA (NOMBRES, TÍTULO, ETC.) %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%
%\fbox{
\begin{minipage}[l][0.78\textheight][t]{0.1\textwidth}
    \begin{center}
    \hskip0pt
    \vrule width2.5pt height17cm    %% height definde la altura de la linea gruesa
        \hskip1mm
        \vrule width1pt height17cm  %% height define la altura de la linea delgada
        \end{center}
\end{minipage}
%}
%\fbox{
\begin{minipage}[c][0.78\textheight][t]{0.8\textwidth}
      \begin{center}
      \vspace{2cm}
        {\Large \scshape {Luminosity measurement with the Pixel detector at the CMS experiment of the Large Hadron Collider}

@@

        \makebox[5cm][c]{\LARGE THESIS}  \\[20pt]
         in partial fulfillment of the
requirements for the degree of:\\[5pt]
        {\Large \textbf{{Doctorado en Ciencias (F\'isica)}}}\\[40pt]
        By:\\[12pt]
        \textbf{  {\large Ashish Sehrawat}}

@@

        { Hermosillo, Sonora} \hskip2.5cm {June, 2023}
      \end{center}
\end{minipage}
%}

### 2023-06-16 tesis_Ashish/chap_originales/Chapter3/chapter3.tex
Your commit message

Figure  \ref{pileup} shows a simulation with pp collision zero-bias events with a full CMS detector response made in 2016, it contain only in time pileup and detector noise, where the mean number of pixel clusters is of the order of  100 per event. For simplicity, the number of pileup interactions present in each event is randomly generated from a representative PCC distribution at $\mu$ = 45 and the average PCC as a function of $\mu$. The latter distribution is fitted with a first-order polynomial, assuming no correlations among different values of $\mu$. Good agreement is seen based on the estimated goodness-of-fit $\chi^{2}$ per degree of freedom (dof) value of about 0.5 , indicating linearity under simulated conditions \cite{lumi_precise_2015_2016}.

\begin{center}
  \begin{figure}[h!]
    \centering
    \includegraphics[scale=.3]{Chapter3/pileup_lineality.png} 
    \caption[Linearity with pile-up]{ The left plot shows the number of pixel clusters and their statistical spread from simulation of pileup following a Poisson distribution with a mean of 45. The right plot shows the mean number of pixel clusters from simulation as a function of mean pileup. The red curve is a first-order polynomial fit with slope and $\chi^{2}/ndof$ values shown in the legend. The lower panel of the right plot shows the difference between the simulation and the linear fit in black points. The green band is the final linearity uncertainty for the 2016 data set \cite{lumi_precise_2015_2016} }
    \label{pileup}
  \end{figure}
\end{center}

@@

This method is capable of providing a precise luminosity measurement over longer time periods, but it cannot do so for a single luminosity section (23 seconds) as is possible with online luminometers. The reason for this limitation is the limited CMS trigger bandwidth available for collecting data \cite{pas_18}. A detailed discussion on this topic can be found in the next chapter.\\

@@

On July 12th, 2022, the data collection for Run3 of the Large Hadron Collider (LHC) began, which resulted in a total luminosity of $38.48 fb^{-1}$. This collection period encompassed the duration from the start of stable beams until the beam was dumped, or when stable beams ceased for other reasons, such as beam studies. The luminosity delivered by the LHC to CMS (blue) and the luminosity recorded by CMS (orange) during stable beams and proton-proton collisions at a center-of-mass energy of $\sqrt{s}$ 13.6 TeV in 2022 are depicted in Figure \ref{Lumi_2022}. Note that  the special runs that occurred at injection energy (450 GeV) are not represented in these graphs.

\begin{center}
  \begin{figure}[h!]
    \centering
    \includegraphics[width=.45\textwidth]{Chapter3/lumi_per_day_cumulative_pp_2022.png}
    \includegraphics[width=.45\textwidth]{Chapter3/lumi_per_day_pp_2022.png}
    \caption[Integrated Luminosity 2022]{(left) Cumulative day-by-day integrated luminosity, 2022 Cumulative measured luminosity versus day delivered. Note that this plot may not include data taken within the most recent 24 hours. (rigth) Day-by-day integrated luminosity, 2022 As the first plot, but not cumulative.} 
    \label{Lumi_2022}
  \end{figure}
\end{center}

@@

To ensure accurate luminosity measurement, a module selection is created to remove any modules exhibiting long-term performance variations, indicating a non-physical shift in their cluster counts. With a total of 1856 modules in the pixel system, all of them can be considered in the luminosity measurement; however, non-zero occupancy and non-linear effects can pose challenges for accurate measurement. Thus, to identify "good" and "stable" modules, a subset is selected by eliminating  underperforming or "bad" modules and comparing their relative contributions to the cross section. Those with relatively consistent contributions across standard physics runs are kept, while those with significant changes in their relative cross section compared to the averaged relative contribution are rejected. The Module selection is achieved  with the following procedure:

\begin{itemize}
\item Poor statistics lumisections are removed by applying selection on total PCC.
\item Barrel layer-1 modules are removed, as these modules are significantly affected by dynamic inefficiency.
\item A loose selection of 7\% based on RMS/mean frome the module weight distribution (changes in the module PCC ratio relative to the total PCC) is applied. Modules with significantly large RMS/mean are removed with this selection as shown in \ref{goodmodule}.
\item The module stability is re-evaluated based on RMS/mean values using an iterative method where the second iteration is made with a loose selection of 4\%.
\end{itemize}

@@

To ensure the accuracy of the measurements, pixel module veto lists were generated for five different time periods, namely C, D, E, F, and G, using the procedure described above for each period. To further enhance the stability of the PCC measurement, a common module vetolist was created by combining the lists from all five periods were the number of bad modules per period were added and modules with equal ID were removed , as shown in table \ref{common veto module}. The PCC data with zero bias was then reprocessed using this common module vetolist and ultimately used for the VdM analysis.

@@

As discussed in Chapter 1, the instantaneous luminosity for a single colliding bunch is described by Eq. (\ref{luminosity_2}). In practice, while the measurement of the beam currents $N_{1,2}$ is well determined, the individual proton density functions cannot be directly measured. To address this, the VdM method involves a specific machine setup that allows for the determination of the two beam overlap integrals. This is achieved by varying the separation between the beams and measuring the resulting rates, which can be used to obtain density profiles that are close to normal distributions.

>>>>

Figure \ref{pileup} shows a simulation with pp collision zero-bias events with the pixel cluster counting rate as a function of pileup in the pileup range observed in Run 2 from 0 to about 50 $\mu$. Where the mean number of pixel clusters is of the order of  111 per event. The red curve is a first-order polynomial fit with slope, a good agreement is seen based on the estimated goodness-of-fit $\chi^{2}$ per degree of freedom (dof) with a value of about 0.5, showing that the PCC rate is highly linear in this range under simulated conditions \cite{ Phase2_Upgrade,lumi_precise_2015_2016}.

\begin{center}
  \begin{figure}[h!]
    \centering
    \includegraphics[scale=.35]{Chapter3/pileup_lineality.png} 
    \caption[Linearity with pile-up]{ Linearity of pixel cluster counting over the pileup range observed in Run 2 (from 0 to about 50) from simulation. The red line is a linear fit to the points \cite{Phase2_Upgrade} }
    \label{pileup}
  \end{figure}
\end{center}

@@

%This method is capable of providing a precise luminosity measurement over longer time periods, but it cannot do so for a single luminosity section (23 seconds) as is possible with online luminometers. The reason for this limitation is the limited CMS trigger bandwidth available for collecting data \cite{pas_18}. A detailed discussion on this topic can be found in the next chapter.\\

%PCC si calcula lumi por cada lumi section.  Esto no tiene que ver con online vs offline.  PCC es un offline luminometro, quiere decir que los datos llegan tarde.  Con PCC solo podemos calcular lumi cada lumi section, debido a la baja estadistica (por la lectura de los datos de zero bias).  otros luminometros tienen precision con NB4 (HF, PLT, BCM1F).

@@

On July 12th, 2022, the data collection for Run3 of the Large Hadron Collider (LHC) began, which resulted in a total recorded luminosity of $38.48 fb^{-1}$. This collection period encompassed the duration from the start of stable beams until the beam was dumped, or when stable beams ceased for other reasons, such as beam studies. The luminosity delivered by the LHC to CMS (blue) and the luminosity recorded by CMS (orange) during stable beams and proton-proton collisions at a center-of-mass energy of $\sqrt{s}$ 13.6 TeV in 2022 are depicted in Figure \ref{Lumi_2022}. Note that  the special runs that occurred at injection energy (450 GeV) are not represented in these graphs \citep{wikicern}.

\begin{center}
  \begin{figure}[h!]
    \centering
    \includegraphics[width=.45\textwidth]{Chapter3/lumi_per_day_cumulative_pp_2022.png}
    \includegraphics[width=.45\textwidth]{Chapter3/lumi_per_day_pp_2022.png}
    \caption[Integrated Luminosity 2022]{(left) Cumulative day by day integrated luminosity. (rigth) Day by day integrated luminosity, 2022 as the first plot, but not cumulative.} 
    \label{Lumi_2022}
  \end{figure}
\end{center}

@@

To ensure accurate luminosity measurement, a module selection is created to remove any modules exhibiting long-term performance variations, indicating a non-physical shift in their cluster counts. With a total of 1856 modules in the pixel system, all of them can be considered in the luminosity measurement; however, instabilities and non-linear effects can pose challenges for accurate measurement. Thus, to identify "good" and "stable" modules, a subset is selected by eliminating  underperforming or "bad" modules and comparing their relative contributions to the cross section. Those with relatively consistent contributions across standard physics runs are kept, while those with significant changes in their relative cross section compared to the averaged relative contribution are rejected. The Module selection is achieved  with the following procedure:

\begin{itemize}
\item Poor statistics lumisections are removed from applying selection on total PCC.
\item Barrel layer-1 modules are removed, as these modules are significantly affected by dynamic inefficiency.
\item A loose selection of 7\% based on RMS/mean frome the module weight distribution (changes in the module PCC ratio relative to the total PCC) is applied. Modules with significantly large RMS/mean are removed with this selection as shown in \ref{goodmodule}.
\item The module stability is re-evaluated based on RMS/mean values using an iterative method where the second iteration is made with a loose selection of 4\%.
\end{itemize}

@@

To ensure the accuracy of the measurements, pixel module veto lists were generated for five different time periods, namely C, D, E, F, and G, using the procedure described above for each period. To further enhance the stability of the PCC measurement, a common module vetolist was created by combining the lists from all five periods were the number of bad modules per period were added and modules with equal ID were removed , as shown in table \ref{common veto module}. The PCC data with zero bias was then reprocessed using this common module vetolist and ultimately used for the vdM analysis.

@@

As discussed in Chapter 1, the instantaneous luminosity for a single colliding bunch is described by Eq. (\ref{luminosity_2}). In practice, while the measurement of the beam currents $N_{1,2}$ is well determined, the individual proton density functions cannot be directly measured. To address this, the vdM method involves a specific machine setup that allows for the determination of the two beam overlap integrals. This is achieved by varying the separation between the beams and measuring the resulting rates, which can be used to obtain density profiles that are close to normal distributions.

### 2023-06-16 tesis_Ashish/chap_originales/Chapter5/chapter5.tex
Your commit message

This work aims to focus on the calibration of luminosity for the PCC  luminometer, based on the van der Meer scans conducted by the CMS Collaboration on November 10 and 11 of 2022 for the fill 8381 of proton-proton collisions at $\sqrt{s}=$ 13.6 TeV, where five scan pairs (3 standard vdM and three BI scans) were analyzed with this method.
A explanation of luminosity importance,the LHC and the CMS experiment, the PCC method as luminometer, and the calibration procedure for estimating $\sigma_{vis}$ were all covered in this work.

This study provides a comprehensive explanation of how data was recorded and procesing, starting from the raw files with cluster reconstruction and processing them with a  selection of  "good" Pixel detector modules. This approach aimed to remove instabilities in luminosity measurement for the corresponding calibration periods,  data also was processed with the  estimation of the background correction to the rates (PCC) from the two super separation periods in this program. The study concludes that the fit obtained for this  estimation is not the best due the background value varied over time for the two SS periods, leading to the decision that the fit model should be changed to one with a constant. All the parameters for this analysis were reported.

The assigned value of the calibration is $\sigma_{vis} = 4163 \pm 3 \text{(stat.)} \pm 12 \text{(syst.)} \text{ mb}$. The analysis and results reported in this work will be useful for a precise luminosity measurement analysis for Run 3 in the CMS experiment of LHC, together with the rest of the luminometers  for a final publication that presents all these final results.

>>>>

This work aims to focus on the calibration of luminosity for the PCC  luminometer, based on the van der Meer scans conducted by the CMS Collaboration on November 10 and 11 of 2022 for the fill 8381 of proton-proton collisions at $\sqrt{s}=$ 13.6 TeV, where five scan pairs (3 standard vdM and three BI scans) were analyzed with this method.
A explanation of luminosity importance,the LHC and the CMS experiment, the PCC method as luminometer, and the calibration procedure for estimating $\sigma_{vis}$ were all covered in this work.\\

This study provides a comprehensive explanation of how data was recorded and procesing, starting from the raw files with cluster reconstruction and processing them with a  selection of  "good" Pixel detector modules. This approach aimed to remove instabilities in luminosity measurement for the corresponding calibration periods,  data also was processed with the  estimation of the background correction to the rates (PCC) from the two super separation periods in this program. The study concludes that the fit obtained for this  estimation is not the best due the background value varied over time for the two SS periods, leading to the decision that the fit model should be changed to one with a variable background level. The variation of the beam overlap integrals and peak rates per BCID are reported in this work..\\

The assigned value of the calibration is $\sigma_{vis} = 4163.3 \pm 3.3 \text{(stat.)} \pm 13 \text{(syst.)} \text{ mb}$. The analysis and results reported in this work will be useful for a precise luminosity measurement analysis for Run 3 in the CMS experiment of LHC, together with the rest of the luminometers  for a final publication that presents all these final results.
